---
title: "dplyr_trial"
author: "Matt Madsen"
date: "September 19, 2016"
output: github_document
---


```{r}
suppressPackageStartupMessages(library(dplyr))
library(gapminder)
```


### str outputs data, shows summary

```{r}
str(gapminder)
```

### shows top of data
```{r}
head(gapminder)
```

### data.frame runs through it all ahhhhhh
```{r}
data.frame(gapminder)
```

### save a data table 
```{r}
gtbl <- tbl_df(gapminder)
gtbl

```

### shows the factors 

```{r}
glimpse(gtbl)
```

### If you feel the urge to store a little snippet of your data:
```{r}
(snippet <- subset(gapminder, country == "Canada"))
```

### Filter, uses logic statements and allows you to find data 

```{r}
filter(gtbl, lifeExp < 29)
filter(gtbl, country == "Rwanda")
```


### Pipe operator

Pipe feds into another replaces brackets/braces

```{r}
gapminder %>% head
```

Use select to subset the data on variables or column 

```{r}
select(gtbl, year, lifeExp)
```

similar but with using pipe operators and only taking top 4

```{r}
gtbl %>%
  select(year, lifeExp) %>%
  head(4)
```


```{r}
gtbl %>%
  filter(country == "Cambodia") %>%
  select(year, lifeExp)
```



