---
title: "hw02_Explore_Gapminder_with_the_tidyverse"
author: "Matt Madsen"
date: "September 20, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explore Gapminder with the tidyverse

####This loads the packages required. Install packages if necessary. 

```{r}
library(tidyverse)
library(gapminder)
```

Then we explore the data frame. 

```{r}
gapminder
str(gapminder)
```


Glimpse gives us a preview. Tibbles makes this sort of obsolete. 

Head is similar, tibble has also made sort of obsolete. 

Tail is better for looking for errors than Head or glimpse, as errors usually occur here. 

```{r}
glimpse(gapminder)
```


Let's filter to just certain rows. Let's select just certain columns. 

```{r}
canada <- filter(gapminder, country == 'Canada' | country == 'Mexico')
filter(gapminder, year > 2000)
filter(gapminder, country %in% c("Bulgaria", "Albania"))
select(gapminder, year, lifeExp)
select(filter(gapminder, country == 'Canada'), year, lifeExp)
filter(select(gapminder, year, lifeExp, country), country == "Canada")
```

Using pipe operator 

ex. main data %>% data manipulation 1 %>% data manipulation 2 

*shortcut command* MAC  CMD SHFT M

Pipeline will autocomplete info in data.frames

```{r}
gapminder %>% 
  filter(country == 'Canada') %>% 
    select(year, lifeExp)

gapminder %>% 
  select(year, lifeExp, country) %>% 
    filter(country == "Canada") %>% 

select(-country)  #minus sign allows to remove. 

gapminder %>% 
  select(starts_with("co"))
```






